---
title: "R Review"
output: html_notebook
---

For this review, we'll be working with dispatch data from the Lincoln Police Department. You can find the data online here: [http://opendata.lincoln.ne.gov/datasets/lpd-dispatch-records](http://opendata.lincoln.ne.gov/datasets/lpd-dispatch-records).

Use the Download menu to select a csv download.

Then move the data file you downloaded into your project folder.

### Load packages

Load the tidyverse, janitor and lubridate packages.

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

### Load data

Load your data into an object called 'dispatches'.

```{r}
dispatches <- read_csv("~/Box/Roper Lab/OnboardingRefresher/LPD_Dispatch_Records.csv")
```

#### Fix column names

Use janitor to make all column names comply with R preferred style - all lowercase and underscores between words.

```{r}
dispatches %>% clean_names()
```


### Analysis questions

#### Datatypes

Look at the documentation for the data. Do all the columns appear to be formatted correctly?

  No, they are not. 

Use lubridate to change the RPT_Date column to ymd format.

```{r}
lubridatedispatches <-dispatches %>% mutate(
  CleanDate= ymd(RPT_Date)
)
```

Find the 13 rows that failed to parse. Why?
  The 13 rows failed to parse because the information was not availabele (?)
  
How many cases are in the data? How many unique cases?

    There are 401,137 cases in this data. (see coding below)

```{r}
lubridatedispatches %>% group_by(OBJECTID) %>% tally()
```
    There are 401,116 unique cases in this data. (see coding below)

```{r}
lubridatedispatches %>% group_by(CASE_) %>% tally()
```

#### Arranging

What are the oldest and most recent cases in the data?

  The most recent case is Case C0-X00014 from March 12, 2020. The oldest case is case B7-000001 from January 1, 2017. See code below 

```{r}
lubridatedispatches %>% arrange(desc(CASE_))
```
```{r}
lubridatedispatches %>% arrange((CASE_))
```

#### Filtering

Create a dataframe called 'missing' with just missing persons cases in it.

```{r}
missing <- lubridatedispatches %>% filter(CFS_LEGEND== "MISSING PERSON")
```

Use the str_detect function to find all the cases that mention O Street in the address.

```{r}
missing %>% filter(str_detect(BLOCK_ADDRESS,"O ST")) 
```

#### Counting

Use the count() function to find the number of dispatches to each neighborhood. 

```{r}
lubridatedispatches %>%
  group_by(NEIGHBOR) %>%
  summarise(
    count = n()
  )
```

Which neighborhood appears the most in the data?

    The actual labeled neighborhood with the most data is downtown. Otherwise, N/A has the most. 

Do you see any limitations of this data when we are counting by neighborhood?

    Not all data is assigned to a neighborhood.

#### Mutating

Create a new column called 'year' that includes just the year from date_fixed.

```{r}
dispatches3 <- lubridatedispatches %>% 
  mutate(year(CleanDate))
```

Use the case_when function to create a new categorical variable dividing the rpt_time column into the 24 hours of the day. Be careful to make sure that each time fits into only one value of the new variable.

```{r}
dispatches3 <- dispatches3 %>%
  mutate(hour = case_when(
    as.numeric(RPT_Time) >= 0 & as.numeric(rpt_time) <= 59 ~ 0,
    as.numeric(rpt_time) >= 100 & as.numeric(rpt_time) <= 159 ~ 1,
    as.numeric(rpt_time) >= 200 &
as.numeric(rpt_time) <= 259 ~ 2,
    as.numeric(rpt_time) >= 300 &
as.numeric(rpt_time) <=359 ~ 3,
    as.numeric(rpt_time) >=400 &
as.numeric(rpt_time) <= 459 ~ 4,
    as.numeric(rpt_time) >= 500 &
as.numeric(rpt_time) <=559 ~ 5,
    as.numeric(rpt_time) >=600 &
as.numeric(rpt_time) <=659 ~ 6,
    as.numeric(rpt_time) >=700 &
as.numeric(rpt_time) <= 759 ~ 7,
    as.numeric(rpt_time) >=800 &
as.numeric(rpt_time) <= 859 ~ 8,
    as.numeric(rpt_time) >= 900 &
as.numeric(rpt_time) <= 959 ~ 9,
    as.numeric(rpt_time) >= 1000 &
as.numeric(rpt_time) <= 1059 ~ 10,
    as.numeric(rpt_time) >= 1100 &
as.numeric(rpt_time) <= 1159 ~ 11,
    as.numeric(rpt_time) >= 1200 &
as.numeric(rpt_time) <= 1259 ~ 12,
    as.numeric(rpt_time) >= 1300 &
as.numeric(rpt_time) <= 1359 ~ 13,
    as.numeric(rpt_time) >= 1400 &
as.numeric(rpt_time) <= 1459 ~ 14,
    as.numeric(rpt_time) >=1500 &
as.numeric(rpt_time) <= 1559 ~ 15,
    as.numeric(rpt_time) >= 1600 &
as.numeric(rpt_time) <= 1659 ~ 16,
    as.numeric(rpt_time) >= 1700 &
as.numeric(rpt_time) <= 1759 ~ 17,
    as.numeric(rpt_time) >= 1800 &
as.numeric(rpt_time) <= 1859 ~ 18,
    as.numeric(rpt_time) >= 1900 &
as.numeric(rpt_time) <= 1959 ~ 19,
    as.numeric(rpt_time) >= 2000 &
as.numeric(rpt_time) <= 2059 ~ 20,
    as.numeric(rpt_time) >= 2100 &
as.numeric(rpt_time) <= 2159 ~ 21,
    as.numeric(rpt_time) >= 2200 &
as.numeric(rpt_time) <= 2259 ~ 22,
    as.numeric(rpt_time) >= 2300 &
as.numeric(rpt_time) <= 2359 ~ 23))

                   
```

What hour of the day sees the most police action?

#### Grouping and summarizing 

How many drug-related dispatches occurred in each year of the data?

    2017- 3300
    2018- 3436
    2019- 3185
    2020- 1252 (code below)

```{r}
lubridatedispatches %>%
  group_by(year(CleanDate)) %>%
  filter(CFS_LEGEND == "DRUGS") %>%
  summarise(
    total=n()
```

Create a new column called month. Then using group_by and summarize, find the maximum, minimum and average number of dispatches per month.

    Maximum numbers of calls per month- 11349
  Minimum number of calls per month- 13
  Average number of calls per month- 9328.767

```{r}
dispatches4 <- dispatches3 %>%
  mutate(
    month = month(CleanDate)
  )
```

The code directly above created Month as a new column. 

```{r}
 dispatches4 %>% 
  mutate(Month = floor_date(CleanDate, "month")) %>% 
  group_by(`Month`) %>% 
    tally () %>% 
  summarise(
    max = max(`n`),
    min = min(`n`),
    avg_dispatch = mean(`n`)
  )
    
  
```
#### Percent change

What was the percent change in total number of dispatches from 2018 to 2019?

  The percent change is .82% (see code below)

```{r}
dispatches4 %>%
  group_by(year(CleanDate)) %>%
  tally()
```

```{r}
((118088-117120)/117120)*100
  
```


#### Line charts

Using ggplot, create a line chart of the number of cases per month and year. Choose an appropriate color for the line, add a title and labels and a choose a theme.

```{r}
dispatches4 %>%
  group_by(year(CleanDate)) %>%
  tally()
ggplot() + geom_line(data=dispatches4, aes(x= CleanDate, y=total))
```

What do you observe about the yearly pattern of police dispatches? Why do you suppose that is?

#### Column charts

Using ggplot, create a column chart that shows the five most common categories in the cfs_legend column. Apply appropriate design decisions to your chart.

```{r}
dispatches4 %>% group_by(CFS_LEGEND) %>% summarise(
total = n()
) %>% top_n(5, total) -> topdispatches
```
```{r}
ggplot() + geom_bar(data=topdispatches, aes(x=CFS_LEGEND, weight=total))
```

